---
title: "DALMKL Discussion"
author: "Kaiqaio Li"
date: "August 31, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Discussion

This package allows us to call the C++ code for DALMKL algorithm efficiently. Here we allowed two loss function for binary classification (logistic loss and hinge loss). The objective function for DALMKL model is as follow:
$$argmin_{\alpha,b}\sum_{i=1}^{N}l\left(y_{i},\sum_{m=1}^{M}K_{m}\alpha_{m}+b\right)+C\sum_{m=1}^{M}\left\Vert \alpha_{m}\right\Vert _{K_{m}},$$
wheras the loss function $$l\left(y,\ f\right)$$ can be written as $$\log\left(1+\exp\left(-yf\right)\right)$$ for logistic loss and $$\left(1-yf\right)_{+}$$ for high loss. To optimize this problem we can use the following R function.

```{r}
library(RMKL)
library(kernlab)
data(benchmark.data)
dd <- data[[1]]
trtrain <- sample(nrow(dd), 140)
k_train <- simplify2array(list(kernelMatrix(rbfdot(sigma = 0.05), 
                dd[trtrain, 1:2]), kernelMatrix(vanilladot(), dd[trtrain, 1:2])))
k_test <- simplify2array(list(kernelMatrix(rbfdot(sigma = 0.05), 
                              dd[-trtrain, 1:2], dd[trtrain, 1:2]), 
                              kernelMatrix(vanilladot(), dd[-trtrain, 1:2], 
                                           dd[trtrain, 1:2])))
y_train <- dd[trtrain, 3]
cri_out <- .01
cri_in <- .000001
maxiter_out <- maxiter_in <- 500
C <- .5
calpha <- 10
mod_hinge <- SpicySVM(k_train, y_train, C, cri_out, cri_in, maxiter_out, maxiter_in,
                      calpha)
alpha_hinge <- mod_hinge$alpha
b_hinge <- mod_hinge$b
prediction_hinge <- predictspicy(alpha_hinge, b_hinge, k_test)

mod_logistic <- SpicyLogit(k_train, y_train, C, cri_out, cri_in, maxiter_out, maxiter_in, 
                           calpha)
alpha_logistic <- mod_logistic$alpha
b_logistic <- mod_logistic$b
prediction_logistic <- predictspicy(alpha_logistic, b_logistic, k_test)
```
Here we can call the C++ implementation for hinge or logistic loss as above. Here the `maxiter_out` and `maxiter_in` are the maximum iteration numbers we tried for the outer and inner cycle. The `cri_out` and `cri_in` are the convergence criterion values in outer and inner cycle. The criterion for inner cycle must be smaller than outer cycle like $0.0001$ ratio and we suggested values as $0.01$ and $0.000001$. The `C` paramter is the tuning parameters, which must be positive, for the block one norm penalty. The `calpha` parameter is proximity parameters using for proximal minimization and also need to be positive. Here we suggest value $10$ for this parameter.
